TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-06-10 paper/draft.tex
P7_tweaks

In this work we build a data-driven model for M dwarf spectra with \thecannon\ (\citealt{Ness:2015}, \citealt{Ho:2017a}, \citealt{Casey:2016}, \citealt{Behmard2019}).
\thecannon\ is a pixel-by-pixel regression for spectral data given a training set of stars with reliable (assumed) accurate labels. The model can both predict a spectrum given labels, and also predict labels given a spectrum. Thus it is a generative model for stellar spectra that can be used for parameter inference.
%AWM: not an expert here, so I won't add anything, but this probably needs more information!

The data-driven approach of \thecannon\ is ideal in certain cases:
if stellar labels are only known for a small number of stars
but there are spectra taken for many more;
if it is computationally expensive to obtain labels for a star, but
there are many stars that need labels;
if there are spectral models that work in one wavelength range or resolution but not in another.
The context of this paper is the last:
there are precise calibrations for metallicities of M dwarf stars in the optical, which have been
used successfully, but (as we note above) there isn't as much success with such modeling in the infrared, and not at \apogee\ resolution.
\thecannon\ bridges this gap: it doesn't require using specific lines; instead it allows us to determine labels from a lot of low level metallicity information present in thousands of lines, and as we demonstrate, does so with very good precision.
%AWM: again, the metallicity metrics Babs/Ryan/Myself use are all in the NIR, and they aren't based on models. The temperature tools use models, but they are calibrated against LBOI measurements. I need to come back to editing later, but this paragraph probably needs to be tweaked a bit.

>>>>

In this work we build a data-driven model for M dwarf spectra with \thecannon\ (\citealt{Ness:2015}, \citealt{Ho:2017a}, \citealt{Casey:2016}, \citealt{Behmard2019}).
\thecannon\ is a pixel-by-pixel regression for spectral data given a training set of stars with reliable (assumed) accurate labels. The model can both predict a spectrum given labels, and also predict labels given a spectrum. Thus it is a generative model for stellar spectra that can be used for parameter inference.
%AWM: not an expert here, so I won't add anything, but this probably needs more information.

The data-driven approach of \thecannon\ is ideal in certain cases:
if stellar labels are only known for a small number of stars
but there are spectra taken for many more;
if it is computationally expensive to obtain labels for a star, but
there are many stars that need labels;
if there are spectral models or techniques that work in one wavelength range or resolution but not in another.
%AWM: seems to me like the context here is all 3. You want to run on ~10,000 stars, which is expensive if you use models. You have labels for a small fraction of these. 
%The context of this paper is the last:
There are empirically-calibrated methods to measure \teff\ and metallicity from flux-calibrated spectra in the optical and NIR, but existing methods to model M dwarf spectra in the NIR at high resolution are computationally expensive, calibrated over a narrow range of \teff\ and metallicity, and/or focus on the $J$-band \citep{Lindgren:2016, Lindgren:2017}. \thecannon\ fills this niche: it doesn't require using specific lines or opacity information that may be missing from the models; instead it allows us to determine labels from a lot of low level metallicity information present in thousands of lines, and as we demonstrate, does so with very good precision.
%AWM: again, the metallicity metrics Babs/Ryan/Myself use are all in the NIR, and they aren't based on models. The temperature tools use models, but they are calibrated against LBOI measurements. I need to come back to editing later, but this paragraph probably needs to be tweaked a bit. 
%AWM: OK, let me know what you think about those changes.
